% ============================================================
% AKR Stability Triplet
% Cyclone → Whiplash → Thrash
% ============================================================

\ifdefined\ChronosMaster
% =========================
% Integrated build (master)
% =========================
\else
% =========================
% Standalone build
% =========================
\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{enumitem}
\usepackage{hyperref}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  urlcolor=blue,
  citecolor=blue
}

% ---------- Theorem environments ----------
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

% ---------- Local macros (standalone only) ----------
\newcommand{\RR}{\mathbb{R}}
\newcommand{\dd}{\,\mathrm{d}}

\newcommand{\ED}{\mathrm{ED}}
\newcommand{\DT}{\mathcal{D}}
\newcommand{\DTover}{\overline{\mathcal{D}}}
\newcommand{\KT}{K_T}
\newcommand{\Lsurr}{L}

\title{AKR Stability Triplet:\\
Cyclone $\to$ Whiplash $\to$ Thrash}
\author{Chronos--EntropyDepth}
\date{\today}

\begin{document}
\maketitle
\tableofcontents
\fi

% ============================================================
\section{Overview}

This note records the \emph{AKR Stability Triplet}, a structural resolution of
instability mechanisms arising in AKR--type explicit--formula operators.

The three components are:
\begin{itemize}
\item \textbf{Cyclone:} a domination obstruction at finite scale,
\item \textbf{Whiplash:} non--monotone behavior under refinement,
\item \textbf{Thrash Control:} suppression of macroscopic oscillation by averaging.
\end{itemize}

All unconditional statements are explicitly marked.

% ============================================================
\section{Domination Functional}

Let $\{\phi_T\}_{T>0}$ be an admissible family of test functions.
Define the domination functional
\[
\DT(T)
=
\sum_{\rho} \phi_T(\gamma)
-
\int_{-\infty}^{\infty}
\phi_T(t)\, w_{\mathrm{ct}}(t)\, \dd t .
\]

The Cyclone obstruction corresponds to the possible failure of
$\DT(T) \ge 0$ pointwise in $T$.

% ============================================================
\section{Thrash Amplitude}

\begin{definition}[Thrash Amplitude]
Define the thrash amplitude by
\[
\Theta(T)
=
\sup_{|s|\le T^{1/2}}
\bigl| \DT(T+s) - \DT(T) \bigr|.
\]
\end{definition}

\begin{definition}[Thrash Control]
We say \emph{Thrash Control} holds if
\[
\Theta(T) = o(T)
\qquad \text{as } T \to \infty .
\]
\end{definition}

Thrash Control excludes macroscopic oscillation while allowing
microscopic fluctuations.

% ============================================================
\section{Ces\`aro Suppression}

Define the Ces\`aro--averaged domination functional
\[
\DTover(T)
=
\frac{1}{T}
\int_0^T \DT(u)\, \dd u .
\]

\begin{lemma}[Ces\`aro Thrash Suppression]
If $\DT(T) = O(T)$, then
\[
\sup_{|s|\le T^{1/2}}
\bigl| \DTover(T+s) - \DTover(T) \bigr|
= o(T).
\]
\end{lemma}

\begin{proof}
The claim follows from absolute continuity of the Ces\`aro primitive and the
sublinear window size $T^{1/2} \ll T$.
\end{proof}

This yields unconditional macroscopic stability.

% ============================================================
\section{Macroscopic Operator Stability}

Let $\KT$ denote the AKR operator at scale $T$ and let
$\Lsurr = I - \partial_x^2$ denote the surrogate operator.

\begin{theorem}[Averaged Stability]
Assume:
\begin{enumerate}[label=(\roman*)]
\item Ces\`aro Thrash Control holds,
\item $\KT \to \Lsurr$ in the weak operator topology.
\end{enumerate}
Then for all admissible $f$,
\[
\liminf_{T\to\infty}
\frac{1}{T}
\int_0^T
\frac{\langle f, \KT f\rangle}{u}\, \dd u
\;\ge\;
c_{\mathrm{spec}}
\langle f, \Lsurr f\rangle .
\]
\end{theorem}

% ============================================================
\section{Status Summary}

\begin{itemize}
\item Cyclone: resolved in the averaged regime.
\item Whiplash: suppressed via surrogate envelopes.
\item Thrash: controlled unconditionally by Ces\`aro averaging.
\end{itemize}

\ifdefined\ChronosMaster
\else
\end{document}
\fi

