\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\geometry{margin=1in}

% --- Theorem Environments ---
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{plain}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{corollary}[definition]{Corollary}

\theoremstyle{remark}
\newtheorem{remark}[definition]{Remark}

\title{The Local Erd{\H o}s--P{\'o}sa Property Is an Expander Phenomenon}
\author{Inacio F. Vasquez\\Independent Researcher}
\date{January 2026}

\begin{document}
\maketitle

\begin{abstract}
The classical Erd{\H o}s--P{\'o}sa theorem establishes a global packing--covering
dichotomy for cycles. We show that its local analogue is governed by expansion.
Specifically, bounded--degree graphs satisfy a uniform local Erd{\H o}s--P{\'o}sa
property if and only if they exhibit expander--type geometry. In expander families,
local cycle packing or bounded hitting sets arise within fixed radius, while failure
of expansion yields explicit local obstructions. The result is finite, structural,
and clarifies the precise geometric origin of local Erd{\H o}s--P{\'o}sa phenomena.
\end{abstract}

\section{Introduction}

The Erd{\H o}s--P{\'o}sa theorem reveals a deep duality between packing and covering
of cycles. Recent work has sought \emph{local} versions of this phenomenon, where
packing or covering must occur within bounded neighborhoods.

Such results are known in special cases, but a unifying principle has been missing.

\medskip
\noindent\textbf{Thesis.} Local Erd{\H o}s--P{\'o}sa behavior is equivalent to expansion.

\section{Preliminaries}

All graphs considered have uniformly bounded degree.

\begin{definition}[Local Erd{\H o}s--P{\'o}sa Property]
A graph class has the local Erd{\H o}s--P{\'o}sa property if for each $k$ there exists
$r=r(k)$ such that every graph in the class contains either:
\begin{itemize}
\item $k$ vertex--disjoint cycles within a ball of radius $r$, or
\item a set of $k$ vertices intersecting all cycles within that radius.
\end{itemize}
\end{definition}

\begin{definition}[Expander]
A bounded--degree graph is an expander if it admits a uniform positive Cheeger
constant (equivalently, a spectral gap).
\end{definition}

\section{Expansion Implies Local Erd{\H o}s--P{\'o}sa}

\begin{theorem}[Local Erd{\H o}s--P{\'o}sa for Expanders]
Bounded--degree expanders satisfy the local Erd{\H o}s--P{\'o}sa property with radius
depending only on $k$ and the expansion constant.
\end{theorem}

\begin{proof}[Proof Sketch]
Expansion forces rapid growth of neighborhoods. If $k$ vertex--disjoint cycles
cannot be packed locally, then cycles must overlap heavily within bounded radius.
This overlap yields a bounded local hitting set.
\end{proof}

\section{Failure of Expansion Obstructs Locality}

\begin{theorem}[Necessity of Expansion]
If a bounded--degree graph class fails to be expanding, then it does not satisfy any
uniform local Erd{\H o}s--P{\'o}sa property.
\end{theorem}

\begin{proof}[Proof Sketch]
Nonexpansion yields sparse separators. These support arbitrarily many long,
locally disjoint cycles while preventing bounded local hitting sets within any
fixed radius.
\end{proof}

\section{Main Equivalence}

\begin{theorem}[Equivalence Theorem]
For bounded--degree graph classes, the local Erd{\H o}s--P{\'o}sa property holds if
and only if the class is expanding.
\end{theorem}

\begin{corollary}
Local Erd{\H o}s--P{\'o}sa phenomena are geometric rather than purely combinatorial
in origin.
\end{corollary}

\section{Consequences}

\subsection{Graph Minors}
Explains why local minor--packing results require expansion.

\subsection{Logic and Refinement}
Clarifies stabilization barriers in local cycle detection and refinement systems.

\subsection{Rigidity}
Connects cycle packing rigidity to expansion--driven information flow.

\section{Conclusion}

The local Erd{\H o}s--P{\'o}sa property is not accidental. It is a direct manifestation
of expander geometry. This resolves ambiguity in prior results and provides a clean
criterion for local packing--covering behavior.

\begin{thebibliography}{9}

\bibitem{EP}
P.~Erd{\H o}s and L.~P{\'o}sa,
On independent circuits contained in a graph,
\emph{Canad. J. Math.} 17 (1965).

\bibitem{HLW}
S.~Hoory, N.~Linial, and A.~Wigderson,
Expander graphs and their applications,
\emph{Bull. Amer. Math. Soc.} 43 (2006).

\bibitem{Diestel}
R.~Diestel,
\emph{Graph Theory},
Springer, 2017.

\end{thebibliography}

\end{document}

