\documentclass[11pt]{article}

% ============================================================
% Chronos Master Guard
% ============================================================
\def\ChronosMaster{}

% ============================================================
% Packages
% ============================================================
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{enumitem}
\usepackage{hyperref}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  urlcolor=blue,
  citecolor=blue
}

% ============================================================
% Theorem Environments
% ============================================================
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

% ============================================================
% Chronos Global Macros
% ============================================================
\input{docs/chronos_macros.tex}

% ============================================================
% Title
% ============================================================
\title{Chronos--EntropyDepth:\\
Structural Limits on Refinement and Computation}
\author{Inacio F. Vasquez}
\date{\today}

\begin{document}
\maketitle
\tableofcontents

% ============================================================
\section{Introduction}

The \emph{Chronos--EntropyDepth} framework studies structural limits on
refinement-based computation. The central invariant, \emph{Entropy Depth}
($\ED$), quantifies the minimum number of admissible refinement steps required
to resolve global information under locality and capacity constraints.

Rather than focusing on algorithmic efficiency, Chronos isolates
\emph{structural obstructions} that prevent rapid global resolution.

% ============================================================
\section{Entropy Depth}

\begin{definition}[Entropy Depth]
Let $\{\mathcal{R}_t\}_{t\ge0}$ be an admissible refinement process acting on a
space of instances with initial entropy $H_0$.
The \emph{entropy depth} $\ED$ is the minimal $t$ such that the residual entropy
falls below a fixed resolution threshold.
\end{definition}

Entropy depth lower bounds express impossibility results for refinement-based
procedures under fixed locality.

% ============================================================
\section{AKR Operators and Explicit Formula Structure}

AKR-type operators arise from explicit-formula representations and encode
interactions between spectral data and arithmetic structure.

Let $\KT$ denote the AKR operator truncated at scale $T$, and let
$\Lsurr = I - \partial_x^2$ denote the canonical surrogate operator.

A central obstruction is the domination inequality
\[
\KT \succeq c\,\Lsurr ,
\]
which may fail pointwise in the truncation parameter $T$.

% ============================================================
\section{Refinement Instabilities}

Two instability mechanisms arise naturally:

\begin{itemize}
\item \textbf{Cyclone:} failure of domination at finite scale.
\item \textbf{Whiplash:} non-monotone behavior of certificates under refinement.
\end{itemize}

These phenomena are structural and persist independently of numerical accuracy.

% ============================================================
\section{Averaging and Stability Mechanisms}

Macroscopic stability is restored by combining:
\begin{itemize}
\item surrogate envelopes that suppress Whiplash,
\item averaging and smoothing that suppress oscillatory Thrash.
\end{itemize}

The resulting stability holds unconditionally at the averaged level.

% ============================================================
\appendix
\section{AKR Stability Triplet: Cyclone--Whiplash--Thrash}
\label{app:akr-stability-triplet}

\input{docs/akr-stability-triplet/akr_stability_triplet.tex}

% ============================================================
\section{Consequences for Entropy Depth}

The AKR Stability Triplet supplies the final structural ingredient needed to
close entropy depth lower bounds within the Chronos framework.

Averaged domination suffices for all Chronos lower-bound arguments; pointwise
domination is required only for conditional sharpenings.

% ============================================================
\section{Status Summary}

\begin{itemize}
\item Cyclone obstruction: resolved in the averaged regime.
\item Whiplash instability: suppressed via surrogate envelopes.
\item Thrash oscillations: controlled unconditionally by Ces\`aro averaging.
\item Remaining gaps: conditional rigidity amplification only.
\end{itemize}

\end{document}

