\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{hyperref}

\geometry{margin=1in}

% --- theorem environments ---
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{plain}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{lemma}[definition]{Lemma}
\theoremstyle{remark}
\newtheorem{remark}[definition]{Remark}

% --- macros ---
\newcommand{\ED}{\mathrm{ED}}
\newcommand{\TC}{\mathrm{TC}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\Hc}{\mathcal{H}}

\title{Finite Capacity and Rigidity:\\
Geometry as Constraint, Forbidden Residuals, and Closure Principles}
\author{Inacio F. Vasquez}
\date{}

\begin{document}
\maketitle

\begin{abstract}
We present a finite-capacity rigidity framework in which geometric and
information-theoretic constraints forbid residual refinement beyond bounded
depth. The formalism isolates capacity as the controlling invariant and yields
closure principles applicable to local refinement systems, entropy-depth
arguments, and admissible verification regimes.
\end{abstract}

\section{Framework}

We work with refinement processes whose evolution is constrained by bounded
information injection per step. The guiding principle is that global progress
cannot exceed local capacity.

\begin{definition}[Refinement Process]
A \emph{refinement process} is a sequence
\[
  \{S_t\}_{t \ge 0}
\]
together with update maps
\[
  S_{t+1} = \Phi_t(S_t, Z_t),
\]
where $Z_t$ is a local transcript satisfying $\I(Z_t) \le C$ for a fixed
capacity bound $C$.
\end{definition}

\section{Capacity and Risk}

\begin{definition}[Risk and Success]
Let $Y$ be a target variable and let $\widehat{Y}$ be an estimator measurable
with respect to a transcript $\sigma(Z_{\le t})$.
The \emph{risk} at time $t$ is
\[
  R_t := \mathbb{E}\bigl[ \ell(\widehat{Y}, Y) \bigr],
\]
for a fixed loss function $\ell$.
\end{definition}

\section{Entropy--Depth Constraint}

\begin{definition}[Entropy--Depth]
The entropy--depth of a refinement process is
\[
  \ED := \sum_{t \ge 0} \Delta H_t,
\]
where $\Delta H_t$ is the entropy reduction achieved at step $t$.
\end{definition}

\begin{theorem}[Capacity Bound]
For any refinement process with per-step capacity bound $C$,
\[
  \ED \;\le\; \TC := \sum_{t \ge 0} \I(Z_t) \;\le\; O(C \cdot T),
\]
where $T$ is the number of refinement steps.
\end{theorem}

\begin{proof}
Each refinement step can reduce entropy by at most the information injected at
that step. Summing over $t$ yields the stated inequality.
\end{proof}

\section{Forbidden Residuals}

\begin{definition}[Forbidden Residual]
A residual structure is \emph{forbidden} if it requires entropy reduction
exceeding $\TC$ under any admissible refinement strategy.
\end{definition}

\begin{lemma}
If a target configuration requires $\ED > \TC$, then it is a forbidden residual.
\end{lemma}

\section{Closure Principle}

\begin{theorem}[Rigidity Closure]
Under bounded capacity, every admissible refinement process stabilizes after
$O(\TC)$ steps. No further semantic progress is possible without violating the
capacity constraint.
\end{theorem}

\begin{remark}
This closure principle underlies entropy-depth lower bounds, locality barriers,
and the impossibility of unbounded adaptive refinement in finite-capacity
systems.
\end{remark}

\section{Conclusion}

Finite capacity enforces rigidity. Geometry, information, and refinement depth
align to forbid residual global structure beyond bounded limits.

\bibliographystyle{plain}
\begin{thebibliography}{9}
\bibitem{shannon}
C. E. Shannon,
\emph{A Mathematical Theory of Communication},
Bell System Technical Journal, 1948.
\end{thebibliography}

\end{document}

