\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}

% Theorem-like environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{corollary}[theorem]{Corollary}

\title{Chronos / EntropyDepth Manuscript}
\author{Inacio F. Vasquez}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
Canonical SAT instantiation (MR-XOR) integrated into Chronos / EntropyDepth.
This manuscript formalizes Mixing Rigidity, Entropy Accumulation, and Hard Transcript Capacity within a concrete SAT gadget framework.
\end{abstract}

\tableofcontents

\section{Introduction}
This manuscript integrates MR-XOR as the canonical SAT instantiation for the Chronos / EntropyDepth program.
It provides a rigorous framework for local-to-global information lower bounds on refinement processes and operator coercivity.

\section{Preliminaries}
Define FO\(^k\)-locality, bounded-degree graphs, refinement processes, and the basic notions of entropy depth and transcript capacity.

\section{Hard Transcript Capacity (HTC)}
Formal statement of HTC for polynomial-time observers, with range-bound justification and capacity arguments.

\section{Mixing Rigidity (MR) and Entropy Accumulation (EA)}
Definitions and lemmas establishing finite local configuration types, configuration pumping, and the EA martingale.

\section{Chronos Lower Bound}
Derivation of
\[
T \ge c\,\frac{H(P_n)}{\log C}
\]
for any admissible FO\(^k\)-local refinement process, with SIC and EA.

\section{MR-Coercivity Bridge}
Finite-defect operator coercivity derived from MR, existence of finite-rank projections \(P\), and bounds on near-kernel dimension.

\section{Canonical SAT Instantiation: Affine XOR Gadgets (MR-XOR)}
\label{sec:mr-xor}

\input{sections/07_mr_xor_sat}

\section{Conclusion}
MR-XOR confirms the Chronos program for a concrete NP-hard family, fully integrating local rigidity, entropy depth, and capacity bounds.

\end{document}

